\sysfull

### 2018-08-05 paper/asplos19/intro.tex

Upon power failures, a device loses the volatile state in its registers, stack, SRAM, and retains the state of any non-volatile memory, such as FRAM. While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave non-volatile state incorrect, partially updated, and may leave checkpointed volatile state inconsistent with non-volatile state. These inconsistencies cause intermittent execution to deviate from continuously-powered behavior, often leading to an unrecoverable application failure~\cite{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for intermittently-powered devices: (i) \emph{software-based programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{hardware-based computer architecture support}~\cite{hicks_isca_2017,idetic,nvp}. Complex architectural changes are expensive to design, verify, and manufacture. New architectures are also inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software approaches are simpler and applicable to existing devices today. \note{Talk about CP-based systems and talk why we focused task-based instead...}Therefore, this work focuses on a key limitation of task-based software approaches, that is, the {\em inflexibility} of statically decomposing a program into tasks.
%
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/intro-figure-horiz.pdf}
    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}). On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
    \label{fig:coalesce}
\end{figure}
%
%\textbf{Task Decomposition of Intermittent Programs.} 
%
\textbf{Crucial Drawback of Static Task Decomposition.} Task-based programming and execution models require a programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to statically decompose a program into a collection of tasks. A \emph{task}, a top level function, can include arbitrary computation that should be executed despite arbitrarily-timed power failures.
The programmer (or a compiler) explicitly expresses task-to-task control flow. Fig.~\ref{fig:coalesce} illustrates how a program's tasks execute and shows how tasks transitions can impose a runtime overhead. At each transition, the system incurs an overhead to track and atomically commit modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}. The more task transitions a program requires, the more commit overhead is incurred by the system at runtime. A programmer may thus create very large tasks in an attempt to reduce task transitions overhead. However, a very large task may require more energy to complete than a device's fixed hardware energy buffer can hold which may lead to a task \emph{non-termination} problem. To eliminate this risk, existing systems require the programmer to decompose a program into relatively small tasks to preserve execution progress. These constraints on task sizing lead to the following dilemma: should large tasks be used, that are efficient but risk non-termination, or small tasks that are guaranteed to complete, but incur a high task transition and commit overhead.

@@

%\textbf{Challenges and Contributions} 
\noindent\note{Carlo's rephrasing} \emph{C1) Task Transition Overhead:} given unpredictable incoming energy, how to save progress as few times as possible? \sys tries to minimize task transition overhead by estimating energy conditions at run-time using \emph{recent execution history} as a metric.
% \noindent\emph{C1) Dynamic Task Sizing:} Given unpredictable energy shots arrival, how to specify the size of a dynamic task on the fly? \sys addresses this challenge by using a novel approach: \textit{recent execution history} as a metric to estimate the environment energy conditions and to specify task sizes. 
%Furthermore, It mitigates the risk of dynamic task re-execution by taking advantage of the certainty that the energy buffer offers and approaches the expected power interrupt carefully.

@@

The rest of this paper is organized as follows. Section~\ref{sec:background} provides background on intermittent computing.
Section~\ref{sec:systemdescription} provides an overview of \sys, while
Section~\ref{sec:task_adaptation} describes \sys's energy-adaptive mechanism. The dynamic group privatization is explained in Section~\ref{sec:memory_virtulaization}. Implementation details of \sys are given in Section~\ref{sec:implementation}. Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's evaluation methodology and results. Section~\ref{sec:related_work} positions \sys in the
context of related work and Section~\ref{sec:conclusions} concludes and discusses
future work.

>>>>

Upon power failures, a device loses the volatile state in its registers, stack, SRAM, and retains the state of any non-volatile memory, such as FRAM. While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave non-volatile state incorrect, partially updated, and may leave checkpointed volatile state inconsistent with non-volatile state. These inconsistencies cause intermittent execution to deviate from continuously-powered behavior, often leading to an unrecoverable application failure~\cite{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for intermittently-powered devices: (i) \emph{software-based programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{hardware-based computer architecture support}~\cite{hicks_isca_2017,idetic,nvp}. Complex hardware architectural changes are expensive to design, verify, and manufacture. New hardware architectures are also inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software approaches are simpler and applicable to existing devices today. Therefore, this work focuses on software approaches. In particular we address the key limitation of task-based programming and execution model, that is the {\em inflexibility} of statically decomposing a program into tasks.
%
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/intro-figure-horiz.pdf}
    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}). On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
    \label{fig:coalesce}
\end{figure}

%\textbf{Task Decomposition of Intermittent Programs.}

\textbf{Crucial Drawback of Static Task Decomposition.} Task-based programming and execution models require a programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to statically decompose a program into a collection of tasks. A \emph{task}, a top level function, can include arbitrary computation that should be executed despite arbitrarily-timed power failures. The programmer (or a compiler) explicitly expresses task-to-task control flow. Fig.~\ref{fig:coalesce} illustrates how a program's tasks execute and shows how tasks transitions can impose a runtime overhead. At each transition, the system incurs an overhead to track and atomically commit modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}. The more task transitions a program requires, the more commit overhead is incurred by the system at runtime. A programmer may thus create very large tasks in an attempt to reduce task transitions overhead. However, a very large task may require more energy to complete than a device's fixed hardware energy buffer can hold which may lead to a task \emph{non-termination} problem. To eliminate this risk, existing systems require the programmer to decompose a program into relatively small tasks to preserve execution progress. These constraints on task sizing lead to the following \emph{dilemma}: should large tasks be used, that are efficient but risk non-termination, or small tasks that are guaranteed to complete but incur a high task transition and commit overhead?

@@

%\textbf{Challenges and Contributions}

\noindent \emph{C1) Task Transition Overhead:} given unpredictable incoming energy, how to save progress as few times as possible? \sys tries to minimize task transition overhead by estimating energy conditions at run-time using \emph{recent execution history} as a metric.
% \noindent\emph{C1) Dynamic Task Sizing:} Given unpredictable energy shots arrival, how to specify the size of a dynamic task on the fly? \sys addresses this challenge by using a novel approach: \textit{recent execution history} as a metric to estimate the environment energy conditions and to specify task sizes. 
%Furthermore, It mitigates the risk of dynamic task re-execution by taking advantage of the certainty that the energy buffer offers and approaches the expected power interrupt carefully.

@@

The rest of this paper is organized as follows. Section~\ref{sec:background} provides background on intermittent computing.
Section~\ref{sec:systemdescription} provides an overview of \sys, while Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism. The dynamic group privatization is explained in Section~\ref{sec:memory_virtulaization}. Implementation details of \sys are given in Section~\ref{sec:implementation}. Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's evaluation methodology and results. Section~\ref{sec:related_work} positions \sys in the context of related work and Section~\ref{sec:conclusions} concludes and discusses future work.
